---
title: "Assignment2"
output: github_document
always_allow_html: true
---

```{r}
library(data.table)
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(dtplyr)
library(Hmisc)
library(leaflet)
library(ggfortify)
```


```{r}
urlfile=("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv")
chs_individual<-read_csv(url(urlfile))

urlfile=("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv")
chs_regional<-read_csv(url(urlfile))

head(chs_individual)
head(chs_regional)
```

#Data Wrangling

1.
```{r}
chs_merged <- 
  chs_individual %>%
  left_join(chs_regional, by="townname")

dim(chs_individual)
dim(chs_regional)
dim(chs_merged)

head(chs_merged)

```

2.
```{r}
chs_merged <-
  chs_merged %>%
  mutate(
    obesity_level = cut(
      bmi,
      breaks = c(0, 14, 22, 24, 100),
      right = F,
      labels = c("Underweight", "Normal", "Overweight", "Obese"),
      ))

chs_merged %>%
  select(obesity_level, bmi)


chs_merged %>% 
  group_by(obesity_level) %>% summarise(max(bmi), min(bmi))
```

3.
```{r}
chs_merged <-
  chs_merged %>%
  mutate(
    smoke_gas_exposure = case_when(smoke==0 & gasstove==0 ~ "None of both",
                                   smoke==1 & gasstove==0 ~ "Smoke only", 
                                   smoke==0 & gasstove==1 ~ "Gas stove only",
                                   smoke==1 & gasstove==1 ~ "Both"
                                         ))
                                
table(chs_merged$smoke_gas_exposure)
```

4.
```{r}
chs_merged %>% 
  group_by(townname) %>% summarise(mean(fev, na.rm=TRUE), sd(fev, na.rm = TRUE), mean(asthma, na.rm = TRUE), sd(asthma, na.rm = TRUE))

chs_merged %>% 
  group_by(male) %>% summarise(mean(fev, na.rm=TRUE), sd(fev, na.rm = TRUE), mean(asthma, na.rm = TRUE), sd(asthma, na.rm = TRUE))

chs_merged %>% 
  group_by(obesity_level) %>% summarise(mean(fev, na.rm=TRUE), sd(fev, na.rm = TRUE), mean(asthma, na.rm = TRUE), sd(asthma, na.rm = TRUE))

chs_merged %>% 
  group_by(smoke_gas_exposure) %>% summarise(mean(fev, na.rm=TRUE), sd(fev, na.rm = TRUE), mean(asthma, na.rm = TRUE), sd(asthma, na.rm = TRUE))

```


#Looking at the data (EDA)
1.
- For each town, the regression line shows that BMI and FEV has positive relationship.
```{r}
chs_merged %>%
  filter(!(townname %in% NA)) %>%
ggplot(mapping = aes(y=bmi, x=fev, fill=townname))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~townname, nrow = 3)
```

2.
- For both histograms, significant relationship between variables were not detected.
```{r}
chs_merged %>%
  filter(!(obesity_level %in% NA)) %>%
ggplot(mapping = aes(x=fev, fill=obesity_level))+
  geom_histogram()
  

chs_merged %>%
  ggplot(mapping = aes(x=fev, fill=smoke_gas_exposure))+
  geom_histogram()+
  scale_fill_viridis_d()

```

3.
- Does not show any relationship.
```{r}
chs_merged %>%
  ggplot(mapping = aes(x=obesity_level, fill=smoke_gas_exposure))+
  geom_bar()
  
```

4.
- The stat summary chart shows that there is a positive relationship between fev and bmi but not between fev and smoke_gas_exposure. 

```{r}

chs_merged %>%
  ggplot()+
  stat_summary(mapping = aes(x=obesity_level, y=fev))
  

chs_merged %>%
  ggplot()+
  stat_summary(mapping = aes(x=smoke_gas_exposure, y=fev))



```

5.
- The leaflet map shows that the red dots, which indicates higher pm2.5 mass, are concentrated in the metroplitan area.
```{r}
pm25_pal = colorNumeric(c('blue','red'), domain = chs_merged$pm25_mass)

leaflet(chs_merged) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircles(lng = ~lon, lat = ~lat, color=~pm25_pal(pm25_mass), opacity=1,fillOpacity=1, radius=500)
```

6.
- regression line shows that there is no relationship between FEV and PM2.5_mass

```{r}
chs_merged %>%
  ggplot(aes(x = pm25_mass, y = fev)) +
  geom_point() +
  geom_smooth(method = "lm")
```

