---
title: "Assignment 4"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(microbenchmark)
library(parallel)
```

# HPC

- Problem 1: Make sure your code is nice

Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  t(apply(mat, 1, cumsum))
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```


- Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():

```{r}
cl <- parallel::makePSOCKcluster(2L)
system.time({
  parallel::clusterSetRNGStream(cl, 1231)
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n = 10000))
  print(mean(ans))
  parallel::stopCluster(cl)
})
```



# SQL

```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

```{r}
dbListTables(con)
```


### Question 1

```{sql, connection=con}
PRAGMA table_info(film)
```

- Below is the table showing the number of movies available in each rating category.

```{sql, connection=con}
SELECT rating, COUNT(*) AS count
FROM film
GROUP by rating
```

### Question 2
- Below is the table showing the average replacement cost and rental rate for each rating category.

```{sql, connection=con}
SELECT rating, AVG(rental_rate) AS avg_rental_rate, AVG(replacement_cost) AS avg_replacement_cost
FROM film
GROUP by rating
ORDER by AVG(rental_rate)
```

### Question 3

```{sql, connection=con}
PRAGMA table_info(film_category)
```

- Below is the table showing the number of films with each category ID.

```{sql, connection=con}
SELECT fc.category_id, COUNT(f.film_id) AS count_film
FROM film AS f 
  INNER JOIN film_category AS fc
  ON f.film_id = fc.film_id
GROUP by fc.category_id
```

### Quention 4

```{sql, connection=con}
PRAGMA table_info(category)
```

- Sports is the name of the most popular category.

```{sql, connection=con}
SELECT fc.category_id, COUNT(f.film_id) AS count_film, c.name
FROM film AS f 
  INNER JOIN film_category AS fc
  ON f.film_id = fc.film_id
  INNER JOIN category AS c
  ON fc.category_id = c.category_id
GROUP by fc.category_id
ORDER by COUNT(f.film_id) DESC
```

### Clean up

```{r}
dbDisconnect(con)
```